### 2019-05-07 styles/texmf/tex/latex/Reply_to_reviewer_template.tex
Update presentation

\documentclass[a4paper,11pt]{article}              %  final version

\usepackage{graphics}
\usepackage{fancyhdr}
\usepackage{amssymb,amsfonts,amsmath,amsthm}
\usepackage{natbib}
\usepackage{bstnotations}
\usepackage[margin=30mm]{geometry}
%\geometry{left=2cm,right=2cm,top=3cm,bottom=3cm}

\usepackage{colortbl}
\usepackage{tikz}
\usepackage{subfigure}

\usepackage{authblk}
\renewcommand\Affilfont{\itshape\small}

\newcommand{\revcomment}[1]{
  \begin{quotation}
    {\small
   \noindent   \textbf{Comment:}
   \em #1
    }    
  \end{quotation}
}

\newcommand{\reply}[1]
{
 \noindent   \textbf{Reply:}
#1}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%                  DOCUMENT
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}
\title{Reply to reviewers\\
	journal\_paper\_ID\\
	\textbf{title}}

\author[1]{A. Author} \author[1]{B. Author}

\affil[1]{Chair of Risk, Safety and Uncertainty Quantification, ETH
  Zurich, Stefano-Franscini-Platz 5, 8093 Zurich, Switzerland}

\date{}
\maketitle

The authors would like to thank the reviewers for their thorough reading
of the manuscript and their constructive comments. Corrections have been
made accordingly and are highlighted in red in the revised version of
the paper.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Reply to reviewer \# 1}
%	\revcomment{The papers fails to cite some published papers that used
%	  enhanced Kriging methods for RBDO. The authors should do better
%	  literature survey.}
%	
%	\reply{In the revised version of the paper, some recent references have
%	  been added such as ...}

\revcomment{Cut/paste from the review each comment}

\reply{when writing a rebuttal, follow the golden rule ``always agree
  with the reviewer (to the largest possible extent), and change as
  little as possible in the paper.''}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Reply to reviewer \# 2}
\revcomment{The papers fails to cite some published papers that used
  enhanced Kriging methods for RBDO. The authors should do better
  literature survey.}

\reply{In the revised version of the paper, some recent references have
  been added such as ...}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Reply to editor}
If anything to be added in this section

%%%%%%%%%%%%%%%%%%%%%%% ADDITIONAL REFERENCES %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% References with bibTeX database:
% e.g.
\nocite{SudretHDR}

\renewcommand{\refname}{Additional references}
\small
\bibliographystyle{chicago}
\bibliography{biblioRSUQ}
\end{document}