### 2019-05-07 styles/texmf/tex/latex/rsuqposter/rsuqposterA0_template.tex

\documentclass{rsuqposterA0}

\author{Ann Author}
\title{A0 Poster template}

\usepackage[framemethod=tikz]{mdframed}
\usepackage{lipsum}

\usepackage{tcolorbox}
\definecolor{mycolor}{rgb}{0.122, 0.435, 0.698}

\newmdenv[innerlinewidth=1pt, roundcorner=1cm,linecolor=mycolor,innerleftmargin=.5cm,
innerrightmargin=.5cm,innertopmargin=6pt,innerbottommargin=6pt]{mybox}

\begin{document}
{	\VERYHuge
some content
}

\colorbox{rsuqlightblue}{
	\begin{minipage}[t][13cm][c]{20cm}
		even bigger
	\end{minipage}
}\hfill
\colorbox{ETH1}{
	\parbox[t][13cm][c]{35cm}{
		this is funny
	}
}
\begin{tcolorbox}[colback=white!5!white,colframe=blue!75!black,title=My nice heading]
\parbox{.3\textwidth}{	This is another \textbf{tcolorbox}}
\end{tcolorbox}

% finally put the footer at the end of the document
\rsuqfooter{}
\end{document}